Example case study in GMSE

### 2021-07-31 notebook/cran_vignettes/SI3.Rmd
Update vignettes SI3 and SI4 for CRAN

To demonstrate how GMSE can be used to understand conflict in social-ecological systems, we build upon the [example of resource management](https://www.biorxiv.org/content/biorxiv/early/2017/11/17/221432.full.pdf#page=5) in the main text. We consider a protected population of waterfowl that exploits and damages agricultural land and is therefore a source of conservation conflict between those that seek the conservation of waterfowl and those that are concerned with the loss of agriculture [e.g., @Fox2017a; @Mason2017; @Tulloch2017; @Cusack2018]. As in the main text example, the objective of the manager is to keep waterfowl at a target abundance that minimises extinction risks, while the objective of farmers is to maximise agricultural production on their landscape. Here we consider a more complex simulation, with a level of detail that more accurately reflects a scenario that might occur in a real social-ecological system where the manager sets policies that incentivise the user to act in a way that ensures the persistence of the resource. The policies are backed up by a manager budget that can be allocated to set costs of actions and thus incentivise users to perform the desired actions. The user also has a budget to carry out actions, and the cost of the user actions is affected by the manager policies and budget. Our objective is not to model the dynamics of a specific system, but to show how GMSE could be parameterised using demographic estimates from empirical studies. We therefore consider an example population in which such estimates are well-reported and readily available.

@@

Our goal is not to provide a detailed case study of the Taiga Bean Geese, but rather to demonstrate how such a case study would be possible in GMSE. For simplicity, here we assess conflict using only the `gmse` function to show how parameter values can be set to provide useful results. Novice R users may prefer to run all of the simulations below using the browser-based GMSE GUI by calling `gmse_gui()` from the R command line. Alternatively, experienced R users may prefer to simulate by [looping time steps](SI2.pdf#page=5) through `gmse_apply`, which allows more flexibility for incorporating custom sub-models and dynamically adjusting parameter values.

Simulations using the default GMSE sub-models described above are run using the `gmse` function, which offers a range of options for setting parameter values (see Table 1 for some select examples). Output of `gmse` is an exhaustive list that includes all resources and observations, all stakeholder decisions and actions, and all landscape properties in each time step of the simulation (see [SI7](SI7.pdf) for a description of key data structures).

@@

Within the default sub-models of GMSE, there are several non-default options that can be set using arguments to `gmse` and `gmse_apply`. Brief explanations of these non-default options appear in the [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) (also [listed](https://confoobio.github.io/gmse/reference/gmse.html) on the [GMSE website](https://confoobio.github.io/gmse/)). Here we further explain some of the less obvious or less intuitive options available in GMSE, for which additional explanation beyond the GMSE documentation may be warranted for application to case studies such as the Taiga Bean Geese case study example demonstrated above. Future versions of GMSE are expected to expand upon these options.

@@

  0. *Density-based observation*. In this option, managers observe resources on a subset of the landscape; subset size is determined by the manager's view as set using the `agent_view` parameter. In practice, this is done by having the manager pick a cell on the landscape at random, then count all of the resources within a distance (in any direction) of `agent_view` cells. Managers then extrapolate the density of resources within this subset to estimate the total number of resources on the landscape. For example, if the manager has an `agent_view` of 1, then they are capable of seeing all of the resources on nine landscape cells (the cell on which they are located, and each adjacent cell). If they count that there are 90 resource on these nine cells, then they calculate a resource density of 10 per cell. They then multiply the density of 10 by the total number of landscape cells to estimate the number of resources on the landscape. The error of this estimate naturally decreases with increasing `agent_view`, such that when the manager is capable of viewing the entire landscape, observation error is zero. If desired, the `times_observe` (default equals 1) option can also be used to allow managers to sample more than one landscape subset in their estimate (e.g., if `times_observe = 2`, then density is estimated from two different locations, with the mean of individual estimates used for a combined estimate of population density). 

@@

The GMSE function `gmse` and its graphical user interface counterpart `gmse_gui` offer wide a suite of options for parameterising simulations to fit empirical case studies of conservation interest using default GMSE natural resource, observation, manager, and user sub-models. Future development of these sub-models might usefully incorporate additional details relevant to specific case studies, such population structure, multiple natural resource and user types, or different manager policy and user action possibilities. [Requests for new features](https://github.com/bradduthie/gmse/issues) and contributions to GMSE can be made through [GitHub](https://github.com/bradduthie/gmse). Additionally, where entirely different types of sub-models are required, the `gmse_apply` function [can be used](https://bradduthie.github.io/gmse_SI2.pdf) to more flexibly simulate social-ecological systems. In [SI4](SI4.pdf), we show an example of this using the same Taiga Bean Geese case study that we did here.

>>>>

To demonstrate how GMSE can be used to understand conflict in social-ecological systems, we build upon the [example of resource management](ms.pdf) in the main text. We consider a protected population of waterfowl that exploits and damages agricultural land and is therefore a source of conservation conflict between those that seek the conservation of waterfowl and those that are concerned with the loss of agriculture [e.g., @Fox2017a; @Mason2017; @Tulloch2017; @Cusack2018]. As in the main text example, the objective of the manager is to keep waterfowl at a target abundance that minimises extinction risks, while the objective of farmers is to maximise agricultural production on their landscape. Here we consider a more complex simulation, with a level of detail that more accurately reflects a scenario that might occur in a real social-ecological system where the manager sets policies that incentivise the user to act in a way that ensures the persistence of the resource. The policies are backed up by a manager budget that can be allocated to set costs of actions and thus incentivise users to perform the desired actions. The user also has a budget to carry out actions, and the cost of the user actions is affected by the manager policies and budget. Our objective is not to model the dynamics of a specific system, but to show how GMSE could be parameterised using demographic estimates from empirical studies. We therefore consider an example population in which such estimates are well-reported and readily available.

@@

Our goal is not to provide a detailed case study of the Taiga Bean Geese, but rather to demonstrate how such a case study would be possible in GMSE. For simplicity, here we assess conflict using only the `gmse` function to show how parameter values can be set to provide useful results. Novice R users may prefer to run all of the simulations below using the browser-based GMSE GUI by calling `gmse_gui()` from the R command line. Alternatively, experienced R users may prefer to simulate by [looping time steps](SI2.pdf) through `gmse_apply`, which allows more flexibility for incorporating custom sub-models and dynamically adjusting parameter values.

Simulations using the default GMSE sub-models described above are run using the `gmse` function, which offers a range of options for setting parameter values (see Table 1 for some select examples). Output of `gmse` is an exhaustive list that includes all resources and observations, all stakeholder decisions and actions, and all landscape properties in each time step of the simulation (see [Default GMSE data structures](SI7.pdf) for a description of key data structures).

@@

Within the default sub-models of GMSE, there are several non-default options that can be set using arguments to `gmse` and `gmse_apply`. Brief explanations of these non-default options appear in the [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) (also listed on the [GMSE website](https://confoobio.github.io/gmse/)). Here we further explain some of the less obvious or less intuitive options available in GMSE, for which additional explanation beyond the GMSE documentation may be warranted for application to case studies such as the Taiga Bean Geese case study example demonstrated above. Future versions of GMSE are expected to expand upon these options.

@@

  0. *Density-based observation*. In this option, managers observe resources on a subset of the landscape; subset size is determined by the manager's view as set using the `agent_view` parameter. In practice, this is done by having the manager count all of the resources within a distance (in any direction) of `agent_view` cells of their location.  Managers then extrapolate the density of resources within this subset to estimate the total number of resources on the landscape. For example, if the manager has an `agent_view` of 1, then they are capable of seeing all of the resources on nine landscape cells (the cell on which they are located, and each adjacent cell). If they count that there are 90 resource on these nine cells, then they calculate a resource density of 10 per cell. They then multiply the density of 10 by the total number of landscape cells to estimate the number of resources on the landscape. The error of this estimate naturally decreases with increasing `agent_view`, such that when the manager is capable of viewing the entire landscape, observation error is zero. If desired, the `times_observe` (default equals 1) option can also be used to allow managers to sample more than one landscape subset in their estimate (e.g., if `times_observe = 2`, then density is estimated from two different locations, with the mean of individual estimates used for a combined estimate of population density). Note that if `times_observe = 1`, then manager estimates might be consistently biased if they are observing from a location with disproportionately many or few resources.

@@

The GMSE function `gmse` and its graphical user interface counterpart `gmse_gui` offer wide a suite of options for parameterising simulations to fit empirical case studies of conservation interest using default GMSE natural resource, observation, manager, and user sub-models. Future development of these sub-models might usefully incorporate additional details relevant to specific case studies, such population structure, multiple natural resource and user types, or different manager policy and user action possibilities. [Requests for new features](https://github.com/bradduthie/gmse/issues) and contributions to GMSE can be made through [GitHub](https://github.com/bradduthie/gmse). Additionally, where entirely different types of sub-models are required, the `gmse_apply` function [can be used](SI2.pdf) to more flexibly simulate social-ecological systems. In [Advanced case study options](SI4.pdf), we show an example of this using the same Taiga Bean Geese case study that we did here.

References
================================================================================
